Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-06 pecomplexity.tex
putting it in Phys Rev E Format

\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{color}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}

\usepackage{pifont} % check and x marks
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\gcc}{{\tt 403.gcc}~}
\newcommand{\svd}{{\tt dgesdd}~}
\newcommand{\svdone}{{\tt dgesdd}$_1$~}
\newcommand{\svdtwo}{{\tt dgesdd$_2$}~}
\newcommand{\svdthree}{{\tt dgesdd$_3$}~}
\newcommand{\svdfour}{{\tt dgesdd$_4$}~}
\newcommand{\svdfive}{{\tt dgesdd$_5$}~}
\newcommand{\svdsix}{{\tt dgesdd$_6$}~}
\newcommand{\naive}{na\"ive}

\parskip=3pt

\newcommand{\row}{{\tt row\_major}~}
\newcommand{\col}{{\tt col\_major}~}

\usepackage{authblk}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newtheorem*{mydef}{Definition}
\title{Determinism, Complexity, and Predictability in Computer Performance\\
Quantifying structured complexity and its role in predictability: With applications to predicting computer performance.
\\ Quantifying Predictability through Structural Complexity \\
{\color{green}Quantifying Predictability through Structural Complexity} 
\\ Quantifying Predictability using Structural Complexity 
\\Decoding Predictability in Complicated Real-Valued Time Series: The balance between complexity and temporal structure
\\Decoding Complicated Time Series: The balance between complexity and temporal structure and the implications for predictability
}

\author[1]{Joshua Garland \thanks{joshua.garland@colorado.edu}}
\author[1]{Ryan James \thanks{ryan.james@colorado.edu}}
\author[1,2]{Elizabeth Bradley \thanks{lizb@colorado.edu}}
\affil[1]{Department of Computer Science\\
  University of Colorado at Boulder\\
  Colorado, USA
}
\affil[2]{Santa Fe Institute\\
  New Mexico, USA
}

@@

\maketitle

\begin{abstract}
[[Joshua: Needs to be rewritten to be more general and use computers as the application not the focus, should be done after we have a first draft of the paper]]  Computers are deterministic dynamical systems \cite{mytkowicz09}.
  Among other things, that implies that one should be able to use
  deterministic forecast rules to predict aspects of their behavior.
  That statement is sometimes---but not always---true. The memory and
  processor loads of some simple programs are easy to predict, for
  example, but those of more-complex programs like \gcc are not.
%%%%%%%%%%%%%%%%%%%%%%
%% I had to change all the \verb|blah| entries because they caused
%% latex to barf if they were in figure captions.  Odd bug.
%%%%%%%%%%%%%%%%%%%%%%
  The goal of this paper is to determine why that is the case. We
  conjecture that, in practice, complexity can effectively overwhelm
  the predictive power of deterministic forecast models. To explore
  that, we build models of a number of performance traces from
  different programs running on different Intel-based computers. We
  then calculate the \emph{permutation entropy}---a temporal entropy
  metric that uses ordinal analysis---of those traces and correlate
  those values against the prediction success.
\end{abstract}

>>>>

\documentclass[%
 %reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
% aps,
%pra,
prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{amssymb}
%\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}

\usepackage{pifont} % check and x marks
\usepackage{color}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\gcc}{{\tt 403.gcc}~}
\newcommand{\svd}{{\tt dgesdd}~}
\newcommand{\svdone}{{\tt dgesdd}$_1$~}
\newcommand{\svdtwo}{{\tt dgesdd$_2$}~}
\newcommand{\svdthree}{{\tt dgesdd$_3$}~}
\newcommand{\svdfour}{{\tt dgesdd$_4$}~}
\newcommand{\svdfive}{{\tt dgesdd$_5$}~}
\newcommand{\svdsix}{{\tt dgesdd$_6$}~}
\newcommand{\naive}{na\"ive}
\newcommand{\row}{{\tt row\_major}~}
\newcommand{\col}{{\tt col\_major}~}
%\newcommand{\color{red}}{{\color[rgb]{1,0,0}}}

%\parskip=3pt

%\usepackage{authblk}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newtheorem*{mydef}{Definition}

%\author[1]{Joshua Garland %\thanks{joshua.garland@colorado.edu}}
%\author[1]{Ryan James \thanks{ryan.james@colorado.edu}}
%\author[1,2]{Elizabeth Bradley \thanks{lizb@colorado.edu}}
%\affil[1]{Department of Computer Science\\
%  University of Colorado at Boulder\\
%  Colorado, USA
%}
%\affil[2]{Santa Fe Institute\\
%  New Mexico, USA
%}

@@

\preprint{APS/123-QED}

%\title{Manuscript Title}% Force line breaks with \\
%\thanks{A footnote to the article title}%

\title{Determinism, Complexity, and Predictability in Computer Performance\\
Quantifying structured complexity and its role in predictability: With applications to predicting computer performance.
\\ Quantifying Predictability through Structural Complexity \\
{\color{green}Quantifying Predictability through Structural Complexity} 
\\ Quantifying Predictability using Structural Complexity 
\\Decoding Predictability in Complicated Real-Valued Time Series: The balance between complexity and temporal structure
\\Decoding Complicated Time Series: The balance between complexity and temporal structure and the implications for predictability
}

\author{Ann Author}
 \altaffiliation[Also at ]{Physics Department, XYZ University.}%Lines break automatically or can be forced with \\
\author{Second Author}%
 \email{Second.Author@institution.edu}
\affiliation{%
 Authors' institution and/or address\\
 This line break forced with \textbackslash\textbackslash
}%

\collaboration{MUSO Collaboration}%\noaffiliation

\author{Charlie Author}
 \homepage{http://www.Second.institution.edu/~Charlie.Author}
\affiliation{
 Second institution and/or address\\
 This line break forced% with \\
}%
\affiliation{
 Third institution, the second for Charlie Author
}%
\author{Delta Author}
\affiliation{%
 Authors' institution and/or address\\
 This line break forced with \textbackslash\textbackslash
}%

\collaboration{CLEO Collaboration}%\noaffiliation

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
[[Joshua: Needs to be rewritten to be more general and use computers as the application not the focus, should be done after we have a first draft of the paper]]  Computers are deterministic dynamical systems \cite{mytkowicz09}.
  Among other things, that implies that one should be able to use
  deterministic forecast rules to predict aspects of their behavior.
  That statement is sometimes---but not always---true. The memory and
  processor loads of some simple programs are easy to predict, for
  example, but those of more-complex programs like \gcc are not.
%%%%%%%%%%%%%%%%%%%%%%
%% I had to change all the \verb|blah| entries because they caused
%% latex to barf if they were in figure captions.  Odd bug.
%%%%%%%%%%%%%%%%%%%%%%
  The goal of this paper is to determine why that is the case. We
  conjecture that, in practice, complexity can effectively overwhelm
  the predictive power of deterministic forecast models. To explore
  that, we build models of a number of performance traces from
  different programs running on different Intel-based computers. We
  then calculate the \emph{permutation entropy}---a temporal entropy
  metric that uses ordinal analysis---of those traces and correlate
  those values against the prediction success.
\begin{description}
\item[Usage]
Secondary publications and information retrieval purposes.
\item[PACS numbers]
May be entered using the \verb+\pacs{#1}+ command.
\item[Structure]
You may use the \texttt{description} environment to structure your abstract;
use the optional argument of the \verb+\item+ command to give the category of each item. 
\end{description}
\end{abstract}

\pacs{Valid PACS appear here}% PACS, the Physics and Astronomy
                             % Classification Scheme.
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
